---
title: "p8105_hw1_WL3011"
author: "Weiqi Liang"
output: html_document
date: "`r Sys.Date()`"
---

# 1. Problem 1

Firstly, the **penguins** dataset and the **palmerpenguins** package are loaded.

```{r setup, results='hide'}
# Load the penguins dataset
library(palmerpenguins)
library(ggplot2)
data("penguins", package = "palmerpenguins")
```

## 1.1 Description of Penguins Dataset

The **penguins** dataset consists of **344** observations about 3 species of penguins: **Adelie**, **Gentoo**, and **Chinstrap**. 

It includes following **8** important variables: `r names(penguins)`.

Besides, it has `r nrow(penguins)` rows and `r ncol(penguins)` columns. All penguins' average flipper length is `r mean(penguins$flipper_length_mm, na.rm = TRUE)` mm.

```{r collapse=TRUE}
names(penguins)
nrow(penguins)
ncol(penguins)
mean(penguins$flipper_length_mm, na.rm = TRUE)
```

## 1.2 Flipper Length vs Bill Length

Using the following code to create a scatterplot of **flipper_length** vs **bill_ length** by **speicies**.

```{r}
# Create scatterplot
p <- ggplot(penguins, aes(x = bill_length_mm, 
                          y = flipper_length_mm,
                          color = species)) +
  geom_point(na.rm = TRUE) +
  labs(title = "Flipper Length vs Bill Length by Species",
       x = "Bill Length (mm)",
       y = "Flipper Length (mm)") +
  theme_bw()

print(p)
```

Next, use **ggsave** to export this scatterplot to the project directory.

```{r}
# save as png
ggsave("penguins_scatterplot.png", plot = p)

```





