---
title: "p8105_hw1_WL3011"
author: "Weiqi Liang"
output: html_document
date: "`r Sys.Date()`"
---

# 1. Problem 1

Firstly, the **penguins** dataset and the **palmerpenguins** package are loaded.

```{r setup, message = FALSE, results='hide'}
# Load the penguins dataset
library(palmerpenguins)
library(ggplot2)
library(dplyr)
data("penguins", package = "palmerpenguins")
```

## 1.1 Description of Penguins Dataset

The **penguins** dataset consists of **344** observations about 3 species of penguins: **Adelie**, **Gentoo**, and **Chinstrap**. 

It includes following **8** important variables: `r names(penguins)`.

Besides, it has `r nrow(penguins)` rows and `r ncol(penguins)` columns. All penguins' average flipper length is `r mean(pull(penguins,flipper_length_mm), na.rm = TRUE)` mm.

```{r collapse=TRUE}
names(penguins)
nrow(penguins)
ncol(penguins)
mean(pull(penguins,flipper_length_mm), na.rm = TRUE)
```

## 1.2 Flipper Length vs Bill Length

Using the following code to create a scatterplot of **flipper_length** vs **bill_ length** by **speicies**.

```{r}
# Create scatterplot
p <- ggplot(penguins, aes(x = bill_length_mm, 
                          y = flipper_length_mm,
                          color = species)) +
  geom_point(na.rm = TRUE) +
  labs(title = "Flipper Length vs Bill Length by Species",
       x = "Bill Length (mm)",
       y = "Flipper Length (mm)") +
  theme_bw()

print(p)
```

Next, use **ggsave** to export this scatterplot to the project directory.

```{r}
# save as png
ggsave("penguins_scatterplot.png", plot = p)

```

# 2. Problem 2

## 2.1 Create Data Frame

Firstly, create a data frame as follows:

```{r message = FALSE}
library(tidyverse)
set.seed(3011)
samp_df = tibble(
  norm_samp = rnorm(10),
  norm_samp_flag = norm_samp > 0,
  char_vector = c("A", "B", "C", "D", "E", "F", "G", "H", "I", "J"),
  factor_vector = factor(rep(c("Level5", "Level2", "Level3"), 
                             length.out = 10))
)
```

```{r}
print(samp_df)
```


## 2.2 Take the Mean of Each Variable

The mean of numeric variables (`norm_samp`) and logical variables (`norm_samp_flag`) will work.
```{r collapse=TRUE}
mean(pull(samp_df, norm_samp)) 
mean(pull(samp_df, norm_samp_flag)) 
```

While the character (`char_vector`) and factor (`factor_vector`) variables cannot have a mean calculated since they are strings instead of numeric or logical variables.
```{r eval=FALSE}
mean(pull(samp_df, char_vector)) 
mean(pull(samp_df, factor_vector)) 
```

## 2.3 Convert Variables 
*   Logical values **TRUE** and **FALSE** are treated as **1** and **0**, respectively. Therefore the mean of logical variables can work.
*   The conversion of character variables will result in **NA** values and a warning message like "NAs introduced by coercion".
*   The conversion of factor variable works, but it may not represent meaningful numeric values. In this case, **Level5**, **Level2**, and **Level3** are converted to numeric values **3**, **1**, and **2** respectively. This means that it converts the ascending order of Level to numeric values.
```{r warning = FALSE, collapse=TRUE}
as.numeric(pull(samp_df,norm_samp_flag))
as.numeric(pull(samp_df,char_vector)) 
as.numeric(pull(samp_df,factor_vector))

```
This helps explain why:

*   Logical variables can be converted into numeric values and their means can be calculated.
*   Character variables cannot be converted to numeric. Therefore the mean of character variables cannot work.
*   Factor variables can be converted into numeric values, however not in a meaningful way. After a forced conversion of ` as.numeric`, the mean can be calculated.






